Thermodynamic response functions in a cell fluid model

### 2024-09-01 main_morse_response_functions.tex
write two equivalent expressions for K_T suitable for two forms of eos

	Thus, in this Section, we presented two forms of the equation of state. The first one expresses the pressure as a function of temperature and chemical potential, $P = P(T, \mu)$, and is represented by Eqs.~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}. The second one expresses the pressure as a function of temperature and density, $P = P(T, \bar{n})$, and is represented by Eqs.~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}. These equations, together with the explicit dependency of density on temperature and chemical potential, $\bar{n} = \bar{n}(T, \mu)$, see Eq.~\eqref{eq:density}, are the basis for the calculation of the thermodynamic response functions in the next Section~\ref{sec:respons_functions}

@@

	\subsection{Isothermal compressibility}
	The isothermal compressibility is defined by
	\begin{equation}
		\label{def:isotherm_compres}
		K_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T,\langle N\rangle}.
	\end{equation}
	Let us perform some transformations to rewrite $K_T$ into a form that is more suitable for the equation of state~\eqref{eq:eosMT}
	\begin{eqnarray*}
		K_T & = & - \frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T,\langle N\rangle}
		\\
		& = & -\rho \left(\frac{\partial \left(\frac{1}{\rho}\right)}{\partial P}\right)_{T,\langle N\rangle}
		\\
		& = & \frac{1}{\rho} \left(\frac{\partial \rho}{\partial P} \right)_{T,\langle N\rangle}
		\\
		& = & \frac{1}{\rho} \frac{\left(\partial \rho / \partial \mu\right)_{T}}
		{\left(\partial P / \partial \mu\right)_{T}},
	\end{eqnarray*}
	where $\rho = \langle N \rangle / V$ is the particle number density. We have omitted the condition of constant $\langle N \rangle$ in the last line of the above equation since we have explicit dependencies on temperature and chemical potential for both pressure $P = P(T, \mu)$, see Eq.~\eqref{eq:eosMT}, and density $\rho = \rho(T, \mu)$, see Eq.~\eqref{eq:density}.
	Applying the Gibbs--Duhem equation
	\begin{equation}
		N{\rm d}\mu = -S{\rm d}T + V{\rm d}P,
	\end{equation}
	at $T = const$ one has 
	\begin{equation*}
		{\rm d} P = \rho {\rm d} \mu,
	\end{equation*}
	or
	\begin{equation}
		\label{eq:rho_dpdm}
		\rho = \left(\frac{\partial P}{\partial \mu}\right)_T.
	\end{equation}
	Substituting this into the last expression for $K_T$, one gets
	\begin{equation}
		K_T = \frac{1}{\rho^2} \left(\frac{\partial \rho}{\partial \mu}\right)_{T}
	\end{equation}
	Finally, from~\eqref{eq:rho_dpdm} it follows that
	\begin{equation}
		\left(\frac{\partial \rho}{\partial \mu}\right)_T = \left(\frac{\partial^2 P}{\partial \mu^2}\right)_T
	\end{equation}
	and ultimately we arrive at the very useful expression for the isothermal compressibility
	\begin{equation}
		K_T = \frac{1}{\rho^2} \left(\frac{\partial^2 P}{\partial \mu^2}\right)_{T}
	\end{equation}

@@

>>>>

	Thus, in this Section, we presented two forms of the equation of state. The first one expresses the pressure as a function of temperature and chemical potential, $P = P(T, \mu)$, and is represented by equivalent Eqs.~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}. The second one expresses the pressure as a function of temperature and density, $P = P(T, \bar{n})$, and is represented by equivalent Eqs.~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}. These equations, together with the explicit dependency of density on temperature and chemical potential, $\bar{n} = \bar{n}(T, \mu)$, see Eq.~\eqref{eq:density}, are the basis for the calculation of the thermodynamic response functions in the next Section~\ref{sec:respons_functions}

@@

	\subsection{Isothermal compressibility}
	The isothermal compressibility is defined by
	\begin{equation}
		\label{def:isotherm_compres}
		K_T = -\frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T,\langle N\rangle}.
	\end{equation}
	Let us perform some transformations to rewrite $K_T$ into a form that is more suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
	\begin{eqnarray*}
		K_T & = & - \frac{1}{V}\left(\frac{\partial V}{\partial P}\right)_{T,\langle N\rangle}
		\\
		& = & -\rho \left(\frac{\partial \left(\frac{1}{\rho}\right)}{\partial P}\right)_{T,\langle N\rangle}
		\\
		& = & \frac{1}{\rho} \left(\frac{\partial \rho}{\partial P} \right)_{T,\langle N\rangle}
		\\
		& = & \frac{1}{\rho} \frac{\left(\partial \rho / \partial \mu\right)_{T}}
		{\left(\partial P / \partial \mu\right)_{T}},
	\end{eqnarray*}
	where $\rho = \langle N \rangle / V$ is the particle number density. We have omitted the condition of constant $\langle N \rangle$ in the last line of the above equation since we have explicit dependencies on temperature and chemical potential for both pressure $P = P(T, \mu)$, see Eq.~\eqref{eq:eosMT}, and density $\rho = \rho(T, \mu)$, see Eq.~\eqref{eq:density}.
	Applying the Gibbs--Duhem equation
	\begin{equation}
		N{\rm d}\mu = -S{\rm d}T + V{\rm d}P,
	\end{equation}
	at $T = const$ one has 
	\begin{equation*}
		{\rm d} P = \rho {\rm d} \mu,
	\end{equation*}
	or
	\begin{equation}
		\label{eq:rho_dpdm}
		\rho = \left(\frac{\partial P}{\partial \mu}\right)_T.
	\end{equation}
	Substituting this into the last expression for $K_T$, one gets
	\begin{equation}
		K_T = \frac{1}{\rho^2} \left(\frac{\partial \rho}{\partial \mu}\right)_{T}
	\end{equation}
	Finally, from~\eqref{eq:rho_dpdm} it follows that
	\begin{equation}
		\left(\frac{\partial \rho}{\partial \mu}\right)_T = \left(\frac{\partial^2 P}{\partial \mu^2}\right)_T
	\end{equation}
	and ultimately we arrive at the very useful expression for the isothermal compressibility
	\begin{equation}
		K_T = \frac{1}{\rho^2} \left(\frac{\partial^2 P}{\partial \mu^2}\right)_{T}
	\end{equation}

@@

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then the most suitable expressions for the isothermal compressibility are
	\begin{equation}
		K_T = \frac{1}{\bar{n}} \left(\frac{\partial P}{\partial \bar{n}}\right)^{-1}_T
	\end{equation}
	and 
	\begin{equation}
		K^*_T = \frac{1}{\bar{n}} \left(\frac{\partial P^*}{\partial \bar{n}}\right)^{-1}_T.
	\end{equation}
